One-dimensional Classical Gas in a Harmonic Trap with Repulsive Interaction

### 2017-03-20 Zhiyu.tex
Mar21 PRE format, tex

\begin{comment}
\subsection{Relaxation in Phase-space}

When describing simple harmonic motion, it is usually convenient to use the language of phase-space (of single particle). In the phase-space, the motion becomes circular, so each of these particles has two independent degrees of freedom, i.e., modulus and phase angle. Here, ``independent" means that the simple harmonic oscillation never ``mix" these two degrees of freedom. In view of this, when it comes to ``thermalization", one may expect that the thermalization of energy, which is just the modulus, does not ensure the thermalization of phase angle. To distinguish these two kind of meanings of ``thermalization", we will call the process of energy distribution going to equilibrium ``thermalization", while the process of phase angle going to equilibrium ``relaxation". The time scale of thermalization and relaxation could be different in principle. We start with the configuration where particles' velocities are set to be zero, which means in phase-space the cloud is line shape at the beginning. If the time scale of thermalization is much longer than that of relaxation, we will see cloud going to isotropic in a short time, and radius distribution slowly evolves to Boltzmann distribution. If the relaxation time is longer than the thermalization time, the cloud will maintain elliptical shape or elliptical shape for a long time while the radius distribution is already Boltzmann.

To study the relaxation, it is convenient to go back to the picture introduced in discussing breathing frequency, since in that chapter, we focus on the average effect of collision on precession motion, but now we focus on the fluctuation near this average effect. Let's recall two distinct regime discussed in Fig.\ref{fig:Breathingfrequency3}. These two regime will show different behavior of relaxation.

In ``bouncing regime", (see Fig.\ref{fig:Breathingfrequency3}(a))particles' position in phase-space have a finite shift $\sim\sigma$ in one collision. As is discussed in former chapter, the average effect of this shift result in precession motion in phase-space and lead to a perturbed breathing frequency. This result is based on approximating the mean effect of collision process with the collision between two particles aligned symmetrically on P axis. However, in reality, the particles 
are everywhere so that the shift do not rotate the particles' phase angle by the same amount in every collision, thus the increase of the phase angle of each particle can be thought of as a wide distribution with a non-zero mean value. As a result of this distribution, the phase angle soon got thermalized, which means the ``relaxation time" is short.

In `` passing regime", (see Fig.\ref{fig:Breathingfrequency3}(b)),the maximal shift of position in phase-space starts to decrease with the decrease of $F_0$. The shift $\delta$ decrease as $\delta\sim F_0^2$. This means the particles are localized in their original position when $F_0$ is smaller-- it is natural to get this conclusion, since the weaker the interaction is, the longer time it takes to ``relax".

All our discussion before is based on the approximation that every particle collide with each other once in one period. But this is hardly true in reality. If the number of collision in every period is taken into account, these two regime looks even more different: In ``bouncing regime", the average precession speed is significant, (see Fig.\ref{fig:Breathingfrequency4}), which means each collision cause a finite precession. The fluctuation of the number of collision one particle has in each period will contribute to the divergence of their phase angles. On the contrary, in ``passing regime", (see Fig.\ref{fig:Breathingfrequency4}), the breathing frequency is close to 2, which means the average effect of collision create zero precession speed. As a result, even though the number of collision is not a constant, it has on average no effect whether one particle collide more or less in one period.

To directly check this argument, we can inspect the frequency spectrum of $R(t)$. As is shown in FIG.\ref{fig:frequencyspectrum}, when $F_0$ is very small (red curve), the peak is very sharp and peak frequency is exactly 2. When $F_0$ goes far beyond the threshold value of $\frac{E}{N\sigma}$ (blue, purple, pink, etc.), the peak becomes a wide distribution with a peak value deviated far from 2, which lead to  fast thermalization according to our argument before.

\begin{figure}[hbtp]
\centering
\includegraphics[scale=0.5]{ZhiyuPictures/freq_spectrum_N=10_E=1000_pre_rev.eps}
\caption{Frequency spectrum, measured over $10^4$ time units.\label{fig:frequencyspectrum}}
\end{figure}

\end{comment}

@@

\section{Shape of Distribution in Phase-space}\label{section:Shape}
%To date, we have verified that the energy distribution will go to Boltzmann distribution. However, even if a system have reached Boltzmann distribution, it doesn't mean that the system has completely ``forgotten" its initial state. There is other information that could still exist, for instance, the shape of distribution cloud in phase-space. One of the reasons why we care about this quantity is that it has a strong effect on breathing behavior. For instance, if the distribution cloud is a thin ellipse (all particles oscillate in phase), the amplitude of oscillation of the cloud radius will be very large. On the contrary, if the cloud is a circle (particles phase of oscillation are completely random), we will not see oscillation at all. Going to equilibrium not only require the thermalization mentioned in last section, but also require reaching a stable configuration in phase-space distribution (relaxation).
When describing simple harmonic motion, it is usually convenient to use the language of phase-space (of single particle). In the phase-space, the motion becomes circular, so each of these particles has two independent degrees of freedom, i.e., modulus and phase angle. Here, ``independent" means that the simple harmonic oscillation never ``mix" these two degrees of freedom. In view of this, when it comes to ``thermalization", one will expect that the thermalization of energy, which is just the modulus, does not ensure the thermalization of phase angle. As ins explained in the Introduction part, in order to distinguish two kinds of meanings of ``thermalization", we will call the process of energy distribution going to equilibrium ``thermalization", while the process of phase angle going to equilibrium ``relaxation". The time scale of thermalization and relaxation could be quite different in principle. We will start with the configuration where particles' velocities are set to be zero, which means in phase-space the cloud is a rotating narrow ellipse at the beginning. If the time scale of thermalization is much longer than that of relaxation, we will see cloud going to some stable shape in a short time (a circle if interaction is not significant compared to total energy) while the distribution of particle number along radius slowly evolves to Boltzmann distribution. If the relaxation time is longer than the thermalization time, the cloud will maintain elliptical shape or some oscillation between different shapes for a long time while the energy distribution is already Boltzmann.

@@

%The time evolution of shape polarization S is shown in Fig.\ref{fig:shape}, when E is much lower than the thermalizaion threshold, S decreases to 0.3 in almost no time. On the contrary, when E is much higher than the threshold value, it takes a long time for S to dacay.

>>>>

@@

\section{Shape of Distribution in Phase-space}\label{section:Shape}
When describing simple harmonic motion, it is usually convenient to use the language of phase-space (of single particle). In the phase-space, the motion becomes circular, so each of these particles has two independent degrees of freedom, i.e., modulus and phase angle. Here, ``independent" means that the simple harmonic oscillation never ``mix" these two degrees of freedom. In view of this, when it comes to ``thermalization", one will expect that the thermalization of energy, which is just the modulus, does not ensure the thermalization of phase angle. As ins explained in the Introduction part, in order to distinguish two kinds of meanings of ``thermalization", we will call the process of energy distribution going to equilibrium ``thermalization", while the process of phase angle going to equilibrium ``relaxation". The time scale of thermalization and relaxation could be quite different in principle. We will start with the configuration where particles' velocities are set to be zero, which means in phase-space the cloud is a rotating narrow ellipse at the beginning. If the time scale of thermalization is much longer than that of relaxation, we will see cloud going to some stable shape in a short time (a circle if interaction is not significant compared to total energy) while the distribution of particle number along radius slowly evolves to Boltzmann distribution. If the relaxation time is longer than the thermalization time, the cloud will maintain elliptical shape or some oscillation between different shapes for a long time while the energy distribution is already Boltzmann.

@@
